---
output:
  html_document:
    keep_md: yes
---
# Reproducible Research: Peer Assessment 1
Margarette K.
```{r, echo=FALSE, results='hide'}
# reset locale to default english
Sys.setlocale(locale = "C")
```

## Loading and preprocessing the data

At first unzip and read our data.
```{r download, echo=TRUE}
unzip("activity.zip", exdir = "data")
activity <- read.csv("data/activity.csv")
```
Now we have dataset called "activity". Formatting dates from factors to dates. Take a look at it with summury function.
```{r summary_activity, echo=TRUE}
activity$date <- as.Date(activity$date)
summary(activity)
```

## What is mean total number of steps taken per day?

Histigram of steps per date.
```{r step_hist, echo=TRUE}
library(ggplot2)
activity.by.date <- aggregate(steps ~ date, data = activity, FUN = sum)
qplot(date, steps, data=activity.by.date, geom = "histogram", stat = "identity", main = "Total number of steps per date", ylab = "number of steps")
```
```{r step_hist_plot_save, echo=FALSE, results='hide'}
ggsave('figures/01_steps_per_date.png', width=4.8, height=4.8, dpi=100)
```

The mean and median total number of steps taken per day.
```{r mean_median, echo=TRUE}
steps.mean <- aggregate(steps ~ date, data = activity, FUN = mean)
names(steps.mean) <- c("date", "mean_steps")
steps.median <- aggregate(steps ~ date, data = activity, FUN = median)
names(steps.median) <- c("date", "median_steps")
merge(steps.mean, steps.median)
```
## What is the average daily activity pattern?

Time series of 5 minute interval on the plot that illustrates activity height during a day.
```{r interval, echo=TRUE}
activity.by.interval <- aggregate(steps ~ interval, data = activity, FUN = mean)
qplot(interval, steps, data=activity.by.interval, geom = "line", main = "Average number of steps per 5-minute interval", xlab = "5-minute interval ", ylab = "number of steps")
```
```{r interval_plot_save, echo=FALSE, results='hide'}
ggsave('figures/02_mean_steps_per_interval.png', width=4.8, height=4.8, dpi=100)
```

The 5-minute interval = 835, on average across all the days in the dataset, contains the maximum number of steps = 206.1698.

```{r max.interval, echo=TRUE}
library(plyr)
interval.max.desc <- arrange(activity.by.interval, desc(steps))
head(interval.max.desc)
```
So, this is the time to go to work 8:35

## Imputing missing values

The total numer of NA in dataset = 2304.
```{r na, echo=TRUE}
sum(is.na(activity))
```
Fill the NA's with the mean for that 5-minute interval.
```{r fill_na, echo=TRUE}
activity.filled <- ldply(lapply(split(activity, as.factor(activity$date)), function(ad) {
  for(i in ad$interval) {
    if (is.na(ad[ad$interval == i, ]$steps)) {
      ad[ad$interval == i, ]$steps <- activity.by.interval[activity.by.interval$interval == i, ]$steps
    }
  }
  ad
}), data.frame)
activity.filled <- activity.filled[, 2:ncol(activity.filled)]
```

Histigram of steps per date.
```{r step_hist_filled, echo=TRUE}
library(ggplot2)
activity.by.date <- aggregate(steps ~ date, data = activity.filled, FUN = sum)
qplot(date, steps, data=activity.by.date, geom = "histogram", stat = "identity", main = "Total number of steps per date", ylab = "number of steps")
```
```{r step_hist_filled_plot_save, echo=FALSE, results='hide'}
ggsave('figures/03_steps_per_date_filled_nas.png', width=4.8, height=4.8, dpi=100)
```

The mean and median total number of steps taken per day.
```{r mean_median_filled, echo=TRUE}
steps.mean <- aggregate(steps ~ date, data = activity.filled, FUN = mean)
names(steps.mean) <- c("date", "mean_steps")
steps.median <- aggregate(steps ~ date, data = activity.filled, FUN = median)
names(steps.median) <- c("date", "median_steps")
merge(steps.mean, steps.median)
```
Now we have medians that are not zero.

## Are there differences in activity patterns between weekdays and weekends?

```{r weekdays, echo=TRUE}
activity.filled$day_type <- as.factor(as.numeric(weekdays(activity.filled$date) %in% c("Saturday", "Sunday")))
levels(activity.filled$day_type)[1] <- 'weekday'
levels(activity.filled$day_type)[2] <- 'weekend'
```

Plotting mean number of steps for weekdays and weekends separately.
```{r weekdays_interval, echo=TRUE}
activity.by.interval <- aggregate(steps ~ interval + day_type, data = activity.filled, FUN = mean)
qplot(interval, steps, data=activity.by.interval, geom = "line", xlab = "5-minute interval ", ylab = "number of steps", facets = day_type ~ .)
```
```{r weekdays_interval_plot_save, echo=FALSE, results='hide'}
ggsave('figures/04_mean_steps_per_interval_weekdays_weekends.png', width=4.8, height=4.8, dpi=100)
```

At weekends the man wakes up late. Activity during a weekend is smoother.